---
title: "Workshop - Subset Selection"
author: "Applied Machine Learning in Economics"
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We are going to use the subset selection on a data set from the [National Longitudinal Survery of Youth 1979 (NLSY79)](https://www.nlsinfo.org/investigator/pages/login) to predict childhood Armed Forces Qualification Test (AFQT) percentile scores.
This is commonly used as a proxy for IQ.

The NLSY has an insane amount of RHS variables.
This workshop is all about selecting the important variables.
We will use forward stepwise selection, backward stepwise selection, and best subset selection on the entire dataset.
Then we will compare the chosen parameters in each model.


# Instructions
***

1. Load the package `leaps` and the dataset.
2. Using `regsubsets` with `nvmax` set to `Inf`, perform forward stepwise selection using the `formula` of `afqt ~ .`
   a. Save the coeficients for the model with the best BIC score
3. Repeat part 2 for backward stepwise selection
4. Repeat part 2 for best subset selection
5. Compare the selections
6. Fit a model with selected coeficients and produce a summary


```{r, include = TRUE}
library(leaps)
nlsy = read.csv('C:/Users/johnj/Documents/Data/Applied ML ECON490/workshop data/nlsy/nlsy_workshop_data.csv')


forw_fit = regsubsets(afqt ~ ., data = nlsy, method = 'forward', nvmax = Inf)
sforw = summary(forw_fit)
coef(forw_fit, which.min(sforw$bic))

back_fit = regsubsets(afqt ~ ., data = nlsy, method = 'backward', nvmax = Inf)
sback = summary(back_fit)
coef(back_fit, which.min(sback$bic))

best_fit = regsubsets(afqt ~ ., data = nlsy, method = 'exhaustive', nvmax = Inf)
sbest = summary(best_fit)
coef(best_fit, which.min(sbest$bic))

fit = lm(afqt ~ glasses_contacts + mother_yos + father_yos + expected_yos + planning +
           working_women_juvenile + race + self_esteem + height + personal_computer + 
           yos + age + life_satisfaction + jobs_num_tot + alcohol,
         data = nlsy)
summary(fit)

cat(names(coef(forw_fit, which.min(sforw$bic))), sep = ' + ')
```

















