---
title: "Random Forest"
author: "Applied Machine Learning in Economics"
date: ""
output: 
  html_document:
    toc: yes
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center')
```

Last time we fit a poorly performing regression tree that predicted the fare paid by titanic passengers.
Today we are going to see if we can do any better.

# Set up
**********
Load

- `tree`
- `randomForest`
- `tidyverse`
- `modelr`

Set the seed to 490.

Load the data as `titanic` 

- select `fare`, `survived`, `pclass`, `sex`, `sibsp`, `age`
- then omit any `NA`s

Finally, sample-split with the training data set to 67%.

```{r, message = FALSE, include = FALSE}
library(tree)
library(randomForest)
library(tidyverse)
library(modelr)
set.seed(490)

titanic = read.csv('C:/Users/johnj/Documents/Data/Applied ML ECON490/workshop data/titanic3.csv')
titanic = titanic %>%
  select(fare, pclass, sex, sibsp, age, survived) %>%
  na.omit

train = sample_frac(titanic, 0.67)
test  = anti_join(titanic, train)
```

Recall we have the MSEs from the previous workshop:

```{r, warning = FALSE}
fit = lm(fare ~ pclass*sibsp*survived, data = train)
mse(fit, test)


fit_tree = tree(fare ~ survived + pclass + sex + sibsp + age, data = train)
fit_best = prune.tree(fit_tree, best = 5)
mse(fit_best, test)
```

# Bagging
*******************************

Using the `formula` of `fare ~ survived + pclass + sex + sibsp + age, data = train`, fit a bagged tree-based model.
specify `na.action = na.omit` in `randomForest()`.
Produce the MSE.

```{r bagging, include = FALSE}
f = fare ~ survived + pclass + sex + sibsp + age
fit_bag = randomForest(f, data = train, mtry = 5, ntree = 490, na.action = na.omit)
fit_bag
mse_bag = mse(fit_bag, test)
mse_bag
```

# Random Forest
*********************
Now fit a random forest with `mtry` set to the square root of the number of covariates.


```{r rf, include = FALSE}
f = fare ~ survived + pclass + sex + sibsp + age
fit_rf = randomForest(f, data = train, mtry = sqrt(5), ntree = 490, na.action = na.omit)
fit_rf
mse_rf = mse(fit_rf, test)
mse_rf
```

Which model has the best performance?
```{r, include = FALSE}
# The random forest is the best performing model
```

