---
title: "Support Vector Machines"
author: "Applied Machine Learning in Economics"
date: ""
output: 
  html_document:
    toc: yes
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We are going to fit a classification SVM on the NLSY data predicting the binary gender variable using height and weight.

Instructions:

1. Load the packages
   - `kernlab`
   - `tidyverse`
2. Set the seed to 490
3. load the NLSY workshop data
4. Perform a 50-50 sample split
5. Using a 5-fold CV with a linear kernel, output the results of a SVM with $C\in\{0.1, 1 \}$ (two models)
6. Using a 5-fold CV with a radial kernel, output the results of a SVM with $C\in\{0.1, 1 \}$ and $\sigma \in \{0.01, 0.1\}$ (four models)
7. Do any of the models stand out as clear winners?
   


```{r, include = FALSE, message = FALSE, warning = FALSE}
library(kernlab)
library(tidyverse)
set.seed(490)
nlsy = read.csv('C:/Users/johnj/Documents/Data/Applied ML ECON490/workshop data/nlsy/nlsy_workshop_data.csv')

train = sample_frac(nlsy, 0.50)
test  = anti_join(nlsy, train)

f = as.factor(msa) ~ mother_yos + yos + I(weight/(height^2)) + firm_size + personal_computer + ever_own_business


(  fitl_0 = ksvm(f, train, C = 0.1, kernel = 'vanilladot', cross = 5)  )
(  fitl_0 = ksvm(f, train, C = 1, kernel = 'vanilladot', cross = 5)  )

(  fitr_0_0 = ksvm(f, train, C = 0.1, kernel = 'rbfdot', kpar = list(sigma = 0.01), cross = 5)  )
(  fitr_1_0 = ksvm(f, train, C = 1, kernel = 'rbfdot', kpar = list(sigma = 0.1), cross = 5)  )
(  fitr_0_1 = ksvm(f, train, C = 0.1, kernel = 'rbfdot', kpar = list(sigma = 0.1), cross = 5)  )
(  fitr_1_1 = ksvm(f, train, C = 1, kernel = 'rbfdot', kpar = list(sigma = 0.1), cross = 5)  )

```
