---
title: "Principal Components"
author: "Applied Machine Learning in Economics"
date: ""
output: 
  html_document:
    toc: yes
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Principal components is a dimension reduction technique.
Today, we are going to identify the number of components that we should use via a scree plot to predict the `self_esteem` variable.
Then we are going to produce a new data set

1. identify the number of components to use via a scree plot
2. create a new dataset from the identified principal components
3. fit a model using `lm()` on the components


Load the NLSY data. No packages or seed setting necessary.

```{r, include = FALSE}
nlsy = read.csv('C:/Users/johnj/Documents/Data/Applied ML ECON490/workshop data/nlsy/nlsy_workshop_data.csv')
```

# Scree Plots
**********************

By first using `scale()`, `cov()`, and `eigen()`, produce a scree plot.

```{r, include = FALSE}
x = model.matrix(self_esteem ~ . - 1, data = nlsy)
pc = eigen(cov(scale(x)))

plot(pc$values)

# It looks like the kink is at 5, so we should use the first 4.
# That is, it looks like the first four principal components hold most the information
```

# New Data
***************

Produce a new dataset using the identified principal components. 
To do so:

1. obtain a matrix of the identified principal components (the eigenvectors) in a matrix called `rotation`
2. Using your rotation matrix, perform the matrix multiplication: `scaled_x %*% rotation` and save the output as `new_x`

```{r, include = FALSE}
rotation = pc$vectors[, 1:4]

new_x = scale(x) %*% rotation
```


# Principal Component Regression
**************************

Fit a PCR model using `lm()` predicting `self_esteem`.
Produce a summary of the fit.

```{r, include = FALSE}
fit_pcr = lm(nlsy$self_esteem ~ new_x)
summary(fit_pcr)
```






