---
title: "Lecture 5: Regression"
author: "ECON 490: Applied ML in Econ"
date: "Fall 2020"
output: 
  powerpoint_presentation:
    reference_doc: ECON_490.potx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
path = 'C:/Users/johnj/Documents/Data/Applied ML ECON490/hw data'
list.files(path)
cps = fread(paste0(path, '/cps_00026.csv'))

names(cps) = cps %>% names %>% tolower # Who wants to hold down shift all the time?
names(cps)

cps = cps[cps$wkswork1 > 40, ]

table(cps$educ)

cps$yos = 0
cps$yos[cps$educ == 1] = NA
cps$yos[cps$educ == 2] = 0
cps$yos[cps$educ == 10] = 4
cps$yos[cps$educ == 20] = 6
cps$yos[cps$educ == 30] = 8
cps$yos[cps$educ == 40] = 9
cps$yos[cps$educ == 50] = 10
cps$yos[cps$educ == 60] = 11
cps$yos[cps$educ %in% c(70:73)] = 12 
cps$yos[cps$educ == 81] = 13
cps$yos[cps$educ %in% c(91,92)] = 14
cps$yos[cps$educ == 111] = 16
cps$yos[cps$educ == 123] = 18
cps$yos[cps$educ %in% c(124:125)] = 20 
cps = na.omit(cps)

cps = cps %>%
  mutate(exp = age - yos - 6,
         female = sex - 1,
         kids = (nchild > 0)*1,
         black = (race == 200)*1)

cps = cps[cps$incwage > 5000, ]


sfit = lm(log(incwage) ~ yos + exp + I(exp^2) + female*kids + black, cps) %>% summary

```

##
$log(wage_i) = log(wage_0) + yos_i \beta_1 + exp_i\beta_2 + exp_i^2 \beta_3 + \epsilon$

$log(wage_i) = log(wage_0) + yos_i \beta_1 + exp_i\beta_2 + exp_i^2 \beta_3 + female_i \beta_4 + kids_i \beta_5 + female_ikids_i\beta_6 + black_i \beta_7 + \epsilon$


```{r, echo = FALSE}
coef(sfit)
a = coef(sfit)[3,1]
b = coef(sfit)[4,1]
summary(cps$exp)

x = seq(0, 40)
y = a + b*x
df = data.frame(x,y)
```



##
```{r, echo = FALSE}

gm = 
  read.csv('C:/Users/johnj/Documents/Data/Applied ML ECON490/lecture data/gapminder_2007.csv')

lm(log(gdpPercap) ~ lifeExp, gm) %>% summary

```


##
$\implies \approx$
```{r, echo = FALSE, message = FALSE}
# Packagages that make me happy
library(tidyverse)
library(data.table)

hw = fread('C:/Users/johnj/Documents/Data/nhis/height_weight.csv') %>% as_tibble
names(hw) = hw %>% names %>% tolower


hw = hw %>% 
  filter(year == 2010,
         height < 95 & height > 0)

male = hw %>%
  filter(sex == 1,
         weight > 126 & weight < 299) %>%
  mutate(Gender = 'Male')

lm(height ~ weight, male) %>% summary
```



##
```{r, echo = FALSE}
x = seq(0,8, length = 1000)
y = (x-4)^2 + 2
df = tibble(x = x, y = y)
ggplot(df, aes(x=x, y=y)) +
  geom_line() +
  labs(x = 'y - yhat', y = 'MSE') +
  theme_classic() +
  ylim(0, 18) +
  theme(axis.ticks = element_blank(), axis.text = element_blank(), text = element_text(size = 20))
```

##
```{r, echo = FALSE}
tibble(x = seq(-2, 8)) %>%
  mutate(y = -.5*x + 3) %>%
  ggplot(aes(x,y)) +
  geom_line() +
  labs(x = 'Education', y = 'dlog(wage)/dEducation') +
  theme_classic() +
  theme(axis.ticks = element_blank(), axis.text = element_blank(), text = element_text(size = 20))
```
## Math

$Y = \alpha + X_1 \beta_1 + ... X_p \beta_p + \epsilon$

$eplison \sim N(0, \sigma^2)$ $R^2$ $\implies$


$\hat{y} = \hat{\alpha} + x_1 \hat{\beta_1} + ... x_p \hat{\beta_p} + e$

$\text{MSE} = \frac{1}{n}\sum_{i=1}^n (y_i - \widehat{f(X_i)})^2$

$$\hat{\alpha}, \hat{\beta} = \text{argmin } \frac{1}{n} \sum_{i=1}^n (y_i - (\alpha + x_1\beta_1))^2$$

## Matrix

$$
\begin{bmatrix}
  y_1\\
  y_2\\
  \vdots\\
  y_n
\end{bmatrix}
=
\begin{bmatrix}
  1\\
  1\\
  \vdots\\
  1
\end{bmatrix}
\alpha + 
\begin{bmatrix}
  x_{1,1}\\
  x_{2,1}\\
  \vdots\\
  x_{n,1}
\end{bmatrix}
\beta_1 + ... +
\begin{bmatrix}
  x_{1,p}\\
  x_{2,p}\\
  \vdots\\
  x_{n,p}
\end{bmatrix}
\beta_p
+ 
\begin{bmatrix}
  e_1\\
  e_2\\
  \vdots\\
  e_n
\end{bmatrix}
$$

$$
\begin{bmatrix}
  y_1\\
  y_2\\
  \vdots\\
  y_n
\end{bmatrix}
=
\begin{bmatrix}
  1 & x_{1,1} & \dots & x_{1,p}\\
  1 & x_{2,1} & \dots & x_{2,p}\\
  \vdots & \vdots & \ddots & \vdots\\
  1 & x_{n,1} & \dots & x_{n,p}
\end{bmatrix}
\begin{bmatrix}
  \alpha\\
  \beta_1\\
  \vdots\\
  \beta_p
\end{bmatrix}
+ 
\begin{bmatrix}
  e_1\\
  e_2\\
  \vdots\\
  e_n
\end{bmatrix}
$$

## Math
$\hat{\beta} = (X^\prime X)^{-1}X^\prime y$

$X_i^2$

$\frac{\partial log(wage)}{\partial Education} = \beta_1 + Education \beta_2$

$$log(Wage) = \alpha + 
\begin{cases}
  \beta_1 & Black = true\\
  0 & Black = flase
\end{cases} +
\begin{cases}
\beta_2 & Female = true\\
\beta_2 & Female = false
\end{cases}
+ \epsilon
$$

## m

$\% \Delta = \frac{New - Old}{Old}$

$f(x) = log(x) \implies f^\prime(x) = \frac{1}{x}$
$\frac{\text{d} f(x)}{\text{d}x} = \frac{1}{x}$
$\text{d} f(x) = \frac{\text{d}x}{x}$

$e^D = \frac{\% \Delta Q}{\% \Delta P}$

$y = e^{\alpha + \epsilon} L^{\beta_1} K^{\beta_2}$
$log(y) = \alpha + log(L)\beta_1 + \log(K)\beta_2 + \epsilon$

## a
$SS_{tot} = \sum_{i=1}^n(y_i - \bar{y})^2$

$\text{V}[y] = \frac{1}{n}SS_{tot}$

$SS_{tot} = SS_{reg} + SS_{res}$

$R^2 = \frac{SS_{tot} - SS_{res}}{SS_{tot}}$
$R^2 = \frac{SS_{reg}}{SS_{tot}}$
$R^2 = \frac{\frac{1}{n}SS_{reg}}{\frac{1}{n}SS_{tot}}$


## h

$H_0: \beta_i = \beta_0$

$H_a: \beta_i \neq \beta_0$


$t = \frac{\hat{\beta_i} - \beta_0}{\text{S}[\hat{\beta_i}]} \sim T(n-(p+1))$

$H_0: \beta_1 = \beta_2 = ... = \beta_p = 0$

$H_a: \beta_j \neq 0 \text{ for at least one } j$

$F = \frac{(SS_{tot} - SS_{res})/p}{RSS/[n - (p + 1)]} \sim F(p, n-(p+1))$

$\text{E}[ \epsilon |X] = 0$ $\text{V}[\epsilon] = \sigma^2$ $cov(\epsilon_i, \epsilon_j) = 0$ $\forall i \neq j$