---
title: "Classification Workshop"
author: "Applied Machine Learning in Economics"
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
In this workshop, we are going to use the [Commodity Flows Survey](https://www.census.gov/programs-surveys/ahs/data/2017/ahs-2017-public-use-file--puf-/ahs-2017-national-public-use-file--puf-.html) to predict export destinations.
See the codebook for details.


# Preliminaries
*****************************

## Setup
****************************

Load in the data as `cfs` and the following packages

- `data.table` (you will actually want this for this data)
- `tidyverse`
- `nnet`
- `MASS`
- `caret`

You don't have to set the variable names to lower case, but I *always* do.

## Trim
*********************

Remove:

1. the export county code "not an export"
2. the standard classification of Transported Goods code "SCTG Suppressed"
3. the mode codes corresponding to "mode suppressed", "single mode", "other mode" and "non-parcel multimode"

## EDA
***********************

Plot a histogram of the shipment value, weight, and distance routed. 
Should we log these?

Do any other EDA that makes you happy!


## Data Cleaning
*******************************

Run this code
```{r, eval = FALSE}
##################################################
## Standard Classification of Transported Goods ##
##################################################

cfs$sctg[cfs$sctg %in% paste(0,as.character(1:5), sep = '')] = '01-05'
cfs$sctg[cfs$sctg %in% paste(0,as.character(6:9), sep = '')] = '06-09'
cfs$sctg[cfs$sctg %in% as.character(10:14)] = '10-14'
cfs$sctg[cfs$sctg %in% as.character(15:19)] = '15-19'
cfs$sctg[cfs$sctg %in% as.character(20:24)] = '20-24'
cfs$sctg[cfs$sctg %in% as.character(25:30)] = '25-30'
cfs$sctg[cfs$sctg %in% as.character(31:34)] = '31-34'
cfs$sctg[cfs$sctg %in% as.character(35:38)] = '35-38'
cfs$sctg[cfs$sctg %in% as.character(39:43)] = '39-43'

############################
## Mode of transportation ##
############################

cfs$mode[cfs$mode %in% c(4,5)] = 3
cfs$mode[cfs$mode %in% c(8:10, 101)] = 7 
cfs$mode[cfs$mode %in% c(15:17)] = 13
```

## Sample-splitting
********************************

Set your seed to 490 and create a train and test data, where the train is 75% of the data.
Anti join by shipment id.


# Analysis
*********************************

We are going to compare the error rate of three different models: multinomial logistic regression, LDA, and QDA.
We are not using KNN because that would take a month to fit (okay, probably not that long but I wouldn't be surprised if it took over an hour).

## Logistic Regression
********************************************

Fit the following simple model:
```{r, eval = FALSE}
mfit = multinom(export_cntry ~ log(shipmt_value) + log(shipmt_wght) + log(shipmt_dist_routed), cfs)
```

Then obtain

1. the summary
2. the p-values for each coefficient
3. the confusion matrix
4. the test error rate

Now run the more complex model:
```{, eval = FALSE}
mfit = multinom(export_cntry ~ log(shipmt_value) + log(shipmt_wght) + log(shipmt_dist_routed) + mode + sctg +                   hazmat + temp_cntl_yn, cfs)
```

Obtain the same four measures.

Which model performs better?
Which export destination do these models have a hard time predicting?

## LDA
*****************************************************

Fit an LDA model to the complex model in the logistic regression.
Obtain 

1. the confusion matrix
2. the test error rate

Does this model have a hard time predicting the same destination as the logistic?
Does it perform better than the logistic?

## QDA
*******************************************
Do the steps in the LDA subsection above for a QDA model.

How does this model do with "that" export destination?
How does it's performance compare?


# The Data Science
*****************************

Which model would you choose and why?
