---
title: "Workshop Sept. 14"
author: "Applied Machine Learning in Economics"
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Kids and Their Candy

Suppose there are $n$ kids, all of whom have random amounts of candy. 
We have some extra candies we can only give to one of them. 
We want to make sure it is fair, so the receiving kid does not receive more than the kid who has the most.

Write a `function` that has the arguments

- `candies`: an arbitrary vecor of kids with different amounts of candy 
- `extra`: the arbitrary amount of extra candy that we have

and for each kid prints out in a complete sentence

- the index of the kid
- if receiving the extra candy will make them have the most of the group

For example, if we have `candies = c(1,2,4,1,3,...)` and `extra = 3`, then giving kid 1 the extras would not exceed the maximum, giving kid 2 would exceed the maximum, and so on.

Here is the starter code:
```{r, results = FALSE}
set.seed(490)

# Candies problem
candies = rpois(5, 4)
extra   = sample(c(2:5), 1) 

candy = function(candies, extra){
  
}

candy(candies, extra)
```

## Maximum Distance

Write a `function` that finds the maximum distance between the elements in two vectors.
There must be one element from each vector.
Have this function print out

- The vector and the index minus the other vector and index that produces that maximum value.

For example, if `a = c(5,7)` and b = `c(0,9)`, then the maxium distance is given by `a[2]` minus `b[1]` with a value of 7.

Here is the starer code:
```{r, results = FALSE}
a = sample(c(0:9), 2)
b = sample(c(0:9), 2)

maxdist = function(a, b){
  
}

maxdist(a,b)
```