---
title: "Tree Based Models"
author: "Applied Machine Learning in Economics"
date: ""
output: 
  html_document:
    toc: yes
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center')
```

We are going to build a regression tree predicting the fare paid by passengers on the Titanic.

# Set up
**********
Load

- `tree`
- `tidyverse`
- `modelr`

Set the seed to 490.

Load the data as `titanic` and sample-split with the training data set to 60%.


```{r, message = FALSE, include = FALSE}
library(tree)
library(tidyverse)
library(modelr)
set.seed(490)

titanic = read.csv('C:/Users/johnj/Documents/Data/Applied ML ECON490/workshop data/titanic3.csv')

train = sample_frac(titanic, 0.60)
test  = anti_join(titanic, train)
```

# Regression Tree Fitting
*************************************

Using the `formula` of `fare ~ survived + pclass + sex + sibsp + age, data = train`

1. fit a regression tree
2. plot it
3. add the relevant text


```{r, message = FALSE, warning = FALSE, include = FALSE}
fit_tree = tree(fare ~ survived + pclass + sex + sibsp + age, data = train)


plot(fit_tree)
text(fit_tree)
```

# Pruning
*******************
Following the lecture notes, identify the optimal size of the regression tree.

Using `ggplot()`, plot the deviance vs size.

```{r, message = FALSE, warning = FALSE, include = FALSE}
fit_tree_cv = cv.tree(fit_tree)

fit_tree_cv
# The optimal size is 3

df = data.frame(size = fit_tree_cv$size, deviance = fit_tree_cv$dev)

ggplot(df, aes(x = size, y = deviance)) +
  geom_point() + 
  geom_line()
```

Refit the pruned tree with the identified size. Plot it with the appropriate text. Obtain the MSE.

```{r, warning = FALSE, include = FALSE}
fit_best = prune.tree(fit_tree, best = 3)

plot(fit_best)
text(fit_best)

mse(fit_best, test)
```

# OLS Comparison

Using the variables identified in the "best" regression tree (making it a fair fight)

1. fit a linear model
2. produce its summary
3. produce the MSE and compare to the regression tree

```{r, include = FALSE}
fit = lm(fare ~ pclass + sibsp, data = train)
summary(fit)
mse(fit, test)

fit = lm(fare ~ pclass*sibsp, data = train)
summary(fit)
mse(fit, test)

# OLS is definitely superior
```









