---
title: "Longitudinal Clustering"
author: "Applied Machine Learning in Economics"
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Setup
*****************

Load the libraries `tidyverse`, `cluster`, and `readxl`. Set the seed to 490. Using `read_excel()`, load the data specifying the argument `sheet = 2`. 

Using one line of code `select()` variables `country`, `year`, `gdp`, and `exports`, omit any NAs, `mutate()` `exports_per_gdp = exports/gdp`, and `filter()` years strictly less than 1885, not `%in%` `1914:1923` and `1940:1947`.



```{r, message = FALSE, warning = FALSE, include = TRUE}
library(tidyverse)
library(cluster)
library(readxl)
set.seed(490)

macro = read_excel('C:/Users/johnj/Documents/Data/macro history/JSTdatasetR4.xlsx', sheet = 2)

mh = macro %>% 
  select(country, year, gdp, exports) %>%
  na.omit %>%
  mutate(exports_per_gdp = exports/gdp) %>%
  filter(year >= 1885 & !year %in% c(1914:1923, 1940:1947))
```

# EDA
***********************

Using `ggplot2`, plot the exports per GDP grouped by country. Which country and year is the maximum?

```{r, include = TRUE}
ggplot(mh, aes(x = year, y = exports_per_gdp, group = country)) +
  geom_line()

mh[which.max(mh$exports_per_gdp), ]
```

# `clusGap()`
**********************

Pivot the data wider, specifying `names_from = year`, `values_from = exports_per_gdp`, and `id_cols = country`.

Using `clusGap()`, specify the clustering function `pam` and `K.max = 8`.
Plot the output using `plot()`.
Determine which is the identified `globalSEmax` number of clusters.

```{r, include = TRUE}
clst = mh %>%
  pivot_wider(names_from = year, values_from = exports_per_gdp, id_cols = country)

cl = clusGap(clst[, -1], FUN = pam, K.max = 8)

plot(cl)

print(cl, method = 'globalSEmax')
```

# Final Plot 
***********************

Refit the model using the identified number of clusters using `pam()`.
Identify the labels for each country.
Reproduce the original plot with colors of exports per GDP over countries colored by their cluster.

Print the labels `arrange()`d by group.

```{r, include = TRUE}
fit = pam(clst[, -1], 4)
lbls = data.frame(country = clst$country, group = fit$clustering)

plt = left_join(mh, lbls)

ggplot(plt, aes(x = year, y = exports_per_gdp, color = as.factor(group), group = country)) +
  geom_line()

lbls %>% arrange(group)
```























