---
title: "GAM"
author: "Applied Machine Learning in Economics"
date: ""
output: 
  html_document:
    toc: yes
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center', fig.width = 16, fig.height = 9)
```

The `geom_smooth()` feature in `ggplot2` using both local regression and splines to produce the smoothed line.
Because local regression is computationally inefficient, `ggplot2` only uses it for data with 1,000 or fewer observations. Otherwise, it uses splines.
We are going to replicate a geom_smooth plot using large number of observations.

Today proceeds in three parts:

1. Loading in the social explorer data and cleaning it
2. Producing a ggplot figure with `geom_smooth`
3. Replicating the figure manually

# Social Explorer Data
**********

These data are from the 2018 ACS 5-year estimates on census block groups retrieved from Social Explorer.
We have the median household income and population density for each of these blocks. 

**Instructions**

1. Load `tidyverse` and `data.table`
2. Load the social explorer data
3. Wrangle and clean the data
   a. Ensure each block has positive population
   b. Define new variables
      i. `id` as the Geo FIPS variable
      ii. `med_house_v` as the block's median house value
      iii. `density` as the block's density
   c. select the three defined variables


```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(data.table)

se = fread('C:/Users/johnj/Documents/Data/Applied ML ECON490/workshop data/social explorer/density vs median house value.csv')

se = se %>%
  filter(SE_A00002_001 > 0) %>%
  mutate(id          = Geo_FIPS,
         med_house_v = SE_A14006_001,
         density     = SE_A00002_002) %>%
  select(id, med_house_v, density)
dim(se)
```

# `geom_smooth()`
**************************

Produce a ggplot with the following:

1. log both variables
2. use `geom_point` with the size set to 0.1
3. use `geom_smooth` with the color set to your favorite, set the standard errors set to false, and set the size to 2

```{r ggplot1, include = FALSE}
ggplot(se, aes(x = log(density), y = log(med_house_v))) +
  geom_point(size = 0.1) + 
  geom_smooth(color = 'cyan', se = FALSE, size = 2)
```

Go through the help documentation for `geom_smooth()`.

1. what type of function is used?
2. what is `"REML"`?
3. what is `"cs"`?

```{r, include = FALSE}
# 1) a spline
# 2) restricted maximum likelihood
# 3) a shrinkage version of the cubic basis spline
```

# DIY `geom_smooth()`
********************
Now it is time to make your own `geom_smooth()`

1. fit the exact same model that `geom_smooth()` uses. 
2. obtain the fitted values as `yhat`
3. obtain the x values as `x`
4. create a data frame `df` that combines `yhat` and `x`
5. produce the same `geom_point()` figure as above 
6. add `geom_line()` with the data set to `df`, the aesthetics set to `x` and `yhat` accordingly, set the size to 2, and set the color 


```{r DIY, include = FALSE}
fit_smooth = mgcv::gam(log(med_house_v) ~ s(log(density), bs = "cs"), data = se,  method = 'REML')
yhat = fit_smooth$fitted.values
x    = fit_smooth$model$`log(density)`
df = data.frame(x = x, yhat = yhat)

ggplot(se, aes(x = log(density), y = log(med_house_v))) +
  geom_point(size = 0.1) + 
  geom_line(data = df, aes(x = x, y = yhat), color = 'cyan', size = 2)
```


